Peer Assessment 1
========================================================

It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the "quantified self" movement -- a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

Loading the activity data file and taking a look at what it contains:

```{r}

temp <- tempfile()
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", method='curl',temp)
activity <- read.table(unz(temp, "activity.csv"), sep=",", header=T)
```

```{r}
names(activity)
str(activity)
```


**What is mean total number of steps taken per day?**

Total mean and median (summary) of number of steps taken per day:
```{r}
summary(aggregate(steps~date,data=activity, FUN=sum, na.rm=T)$steps)
```
Mean and Median seperately:
```{r}
mean(aggregate(steps~date,data=activity, FUN=sum, na.rm=T)$steps)
median(aggregate(steps~date,data=activity, FUN=sum, na.rm=T)$steps)
```

Histogram of the total number of steps taken each day:
```{r fig.width=7, fig.height=6}
hist(aggregate(steps~date,data=activity, FUN=sum, na.rm=T)$steps, 
     xlab="Steps Per Day", 
     main='Total Number of Steps Taken per Day',
     breaks=20)
```

**What is the average daily activity pattern?**

Time series plot of the 5-minute interval and the average number of steps taken, averaged across all days:
```{r}
mean_steps <- aggregate(steps ~ interval, data=activity, FUN = mean, na.rm = TRUE)
plot(mean_steps, type="l", lwd = 5)
```

**Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?**
```{r}
maxNumberSteps <- mean_steps[mean_steps$steps== max(mean_steps$steps),]
print(paste('The maximum number of steps is', round(maxNumberSteps$steps[[1]]),'at interval',maxNumberSteps$interval[[1]]))
```

**Imputing missing values**

Total number of missing values in the dataset:
```{r}
print(paste(sum(is.na(activity)), 'rows with NAs found'))
```

**Devise a strategy for filling in all of the missing values in the dataset. The strategy does 
not need to be sophisticated. For example, you could use the mean/median for that day, 
or the mean for that 5-minute interval, etc.**

Taking the mean of each interval for imputation:
```{r}
meanPerInterval <- aggregate(steps~interval,data=activity, FUN=mean, na.rm=T)
activityImputed <- activity
activityImputed$steps <- ifelse(is.na(activityImputed$steps), 
                    meanPerInterval$steps[match(activityImputed$interval,meanPerInterval$interval)], 
                    activityImputed$steps)
```

** Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.**
```{r}
hist(aggregate(steps~date,data=activityImputed, FUN=sum, na.rm=T)$steps, 
     xlab="Steps Per Day",
     main='Total Number of Steps Taken per Day - Imputed',
     breaks=20)
```

** Do these values differ from the estimates from the first part of the assignment? **

Summary of activity data set:
```{r}
summary(activityImputed$steps)
```

Summary of imputed activity data set:
```{r}
summary(activity$steps)
```

By imputing the data set, the median has shifted upwards from 1065 to 1066 while the mean remained unchanged:
```{r}
mean(aggregate(steps~date,data=activity, FUN=sum, na.rm=T)$steps)
median(aggregate(steps~date,data=activity, FUN=sum, na.rm=T)$steps)
mean(aggregate(steps~date,data=activityImputed, FUN=sum, na.rm=T)$steps)
median(aggregate(steps~date,data=activityImputed, FUN=sum, na.rm=T)$steps)
```
 
**Are there differences in activity patterns between weekdays and weekends?**

Adding the variable 'weektype' to the imputed data set:
```{r}
activityImputed$weektype <- as.factor(ifelse(weekdays(as.Date(activityImputed$date, "%Y-%m-%d")) %in% c('Sunday','Saturday'),
                            'weekend',
                            'weekday'))
```

Panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, 
averaged across all weekday days or weekend days:
```{r}
par(mfrow=c(2,1))
plot(aggregate(steps~interval, 
               data=activityImputed[(activityImputed$weektype =='weekend'),],FUN=mean),
     type='l',
     main='Weekend',
     lwd=2,
     col='blue')
plot(aggregate(steps~interval, 
               data=activityImputed[(activityImputed$weektype =='weekday'),],FUN=mean),
     type='l',
     main='Weekday',
     lwd=2,
     col='red')
```

